---
title: "Clustering Visualization"
output: html_document
date: "2024-05-29"
---

```{r, warnings = FALSE, echo = FALSE, include=FALSE}
library(caret)
library(survival)
library(survminer)
library(ggplot2)
library(randomForest)
library(neuralnet)
library(shapviz)
library(kernelshap)
```

```{r}
data <- read.csv("SEER_final.csv")[,-1]
```

Setting scaling factors and scaling SEER data for clustering
```{r}
set.seed(1)
all <- subset(data, Stage == "Localized")
rownames(all) <- 1:nrow(all)
all.clust <- all[,c(1:9)]
all.save <- all[as.numeric(rownames(all.clust)),]

preProcValues_clust <- preProcess(all.clust, method = c("center", "scale"))

all.clust <- predict(preProcValues_clust, all.clust)
```

```{r}
set.seed(2)

gmm = GMM(all.clust, 25, dist_mode = "eucl_dist", seed_mode = "random_subset", km_iter = 10,
          em_iter = 10, verbose = F)   
pr = predict(gmm, newdata = all.clust)
```

```{r}
#opt_gmm_bic = Optimal_Clusters_GMM(all.clust, max_clusters = 100, criterion = "BIC", 
#                               
#                               dist_mode = "eucl_dist", seed_mode = "random_subset",
#                               
#                               km_iter = 10, em_iter = 10, var_floor = 1e-10)
```

```{r}
#bic_df <- data.frame("Clusters" = 1:100, "BIC" = opt_gmm_bic)
#bic_plot <- ggplot(bic_df, aes(x = Clusters, y = BIC)) +
#  geom_point() +
#  geom_line() +
#  theme_bw() +
#  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 14), axis.title = element_text(family = "serif", size = 14), axis.text = element_text(family = "serif", size = 12))
#bic_plot
```


```{r}
library(Rtsne)
rownames(all.clust) <- 1:nrow(all.clust)
unique_all.clust <- unique(all.clust)
tsne_res <- Rtsne(unique_all.clust)
tsne_vals <- as.data.frame(tsne_res$Y)
colnames(tsne_vals) <- c("tSNE1", "tSNE2")
tsne_vals$Cluster <- as.factor(pr[as.numeric(rownames(unique_all.clust))])
tsne_vals$Breslow_Depth <- all$Breslow_Depth[as.numeric(rownames(unique_all.clust))]
tsne_vals$Rows <- rownames(unique_all.clust)
```


```{r}
p_tsne <- ggplot(subset(tsne_vals), aes(x=tSNE1, y=tSNE2, color=Cluster))+
  geom_point(size = 0.3)+
    theme_bw()+
  theme(plot.title=element_text(hjust=0.5, family = "serif", size = 26), axis.title = element_text(family = "serif", size = 30), axis.text = element_text(family = "serif", size = 26), legend.key.size = unit(0.1, 'cm'), legend.title = element_text(hjust=0.5, family = "serif", size = 30), legend.text = element_text(family = "serif", size = 26)) 
print(p_tsne)
```
```{r}
bd_tsne <- ggplot(subset(tsne_vals), aes(x=tSNE1, y=tSNE2, color=Breslow_Depth))+
  geom_point(size = 0.3)+
    theme_bw()+
  theme(plot.title=element_text(hjust=0.5, family = "serif", size = 26), axis.title = element_text(family = "serif", size = 30), axis.text = element_text(family = "serif", size = 26), legend.key.size = unit(0.1, 'cm'), legend.title = element_text(hjust=0.5, family = "serif", size = 30), legend.text = element_text(family = "serif", size = 26), legend.key.height = unit(5, "null"))  +
  scale_color_gradient(low="#26C6DA", high="#ef5350")
print(bd_tsne)
```

```{r}
set.seed(1)
all.save$Group <- pr

calc_sig <- function(i, data){
  data$Comparison <- NA
  data$Comparison[data$Cluster != i] <- "NC"
  data$Comparison[data$Cluster == i] <- "C"
  if (length(unique(data$Comparison)) > 1){
  logrank <- survdiff(Surv(MSS_Time, MSS) ~ Comparison, data = data)
  pval <- logrank$pvalue
  if ((logrank$obs[[1]]/nrow(subset(data, Comparison == "C"))) > (logrank$obs[[2]]/nrow(subset(data, Comparison == "NC")))){
    risk <- "HR"
  }
  if ((logrank$obs[[1]]/nrow(subset(data, Comparison == "C"))) < (logrank$obs[[2]]/nrow(subset(data, Comparison == "NC")))){
  risk <- "LR"
  }
  if ((logrank$obs[[1]]/nrow(subset(data, Comparison == "C"))) == (logrank$obs[[2]]/nrow(subset(data, Comparison == "NC")))){
  risk <- NA
  }
  if (pval < 0.01){
    sig <- "S"
  }
  if (pval >= 0.01){
  sig <- "NS"
  }
  if (logrank$obs[[1]] == 0){
    z <- 1
  }
    if (logrank$obs[[1]] > 0){
    z <- 0
    }
  surv <- nrow(subset(data, Cluster == i & MSS == 1 & MSS_Time <= 60))/nrow(subset(data, Cluster == i))
  row <- data.frame("Risk" = risk, "Significance" = sig, "Cluster" = i, "Z" = z, "Surv" = surv)
  return(row)
  }
}

```

```{r}
set.seed(1)
all$Cluster <- as.factor(pr)
bd1 <- subset(all, Breslow_Depth <= 0.5 & Stage == "Localized")

data2 <- data.table::rbindlist(lapply(1:25, calc_sig, bd1))
```

```{r}
bd1$Risk <- "NS"
bd1$Risk[bd1$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd1$Risk[bd1$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

#bd1.tsne <- tsne_vals[rownames(subset(unique_all.clust, Breslow_Depth <= 0.5)),]
bd1.tsne <- subset(tsne_vals, Breslow_Depth <= 0.5)

bd1.tsne$Risk <- "NS"
bd1.tsne$Risk[bd1.tsne$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd1.tsne$Risk[bd1.tsne$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

```

```{r}
bd1.clusts <- ggplot(bd1.tsne, aes(x=tSNE1, y=tSNE2, color=Cluster))+
  geom_point(size = 0.1)+
    theme_bw()+
  ggtitle("Breslow Depth \u2264 0.5 mm") +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd1.clusts
```

```{r}

bd1.tsne$Risk <- factor(bd1.tsne$Risk, levels = c("High-Risk", "Low-Risk", "NS"))
bd1.risk <- ggplot(bd1.tsne, aes(x=tSNE1, y=tSNE2, color=Risk))+
  geom_point(size = 0.1)+
  ggtitle("Breslow Depth \u2264 0.5 mm") +
    theme_bw()+
  scale_color_manual(values = c("#ef5350", "#26C6DA", "#B0BEC5")) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd1.risk
```

```{r}
logrank <- survdiff(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd1, Risk != "NS"))
logrank
```

```{r}
pval_label <- expression(paste(italic("p"), " < 0.0001"))
```


```{r}
fit.clust <- survfit(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd1, Risk != "NS"))

bd1_surv <- ggsurvplot(fit = fit.clust, data = subset(bd1, Risk != "NS"), censor.size = 1)$plot
bd1_surv <- bd1_surv +
  ggtitle("Breslow Depth \u2264 0.5 mm") +
    theme_bw()+
    xlab("Months") +
    ylab("Survival Probability")+
  scale_color_manual(values = c("#ef5350", "#26C6DA")) +
  annotate(geom="text", x=25, y=0.2, label=pval_label,
              color="black", size = 8) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22), legend.text = element_text(family = "serif", size = 22), legend.title = element_text(family = "serif", size = 22))
bd1_surv
```
########### 0.5 - 1 mm 

```{r}
set.seed(1)
all$Cluster <- as.factor(pr)
bd2 <- subset(all, Breslow_Depth > 0.5 & Breslow_Depth <= 1 & Stage == "Localized")

data2 <- data.table::rbindlist(lapply(1:25, calc_sig, bd2))
```

```{r}
bd2$Risk <- "NS"
bd2$Risk[bd2$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd2$Risk[bd2$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

bd2.tsne <- subset(tsne_vals,Breslow_Depth > 0.5 & Breslow_Depth <= 1)

bd2.tsne$Risk <- "NS"
bd2.tsne$Risk[bd2.tsne$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd2.tsne$Risk[bd2.tsne$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

```

```{r}
bd2.clusts <- ggplot(bd2.tsne, aes(x=tSNE1, y=tSNE2, color=Cluster))+
  geom_point(size = 0.1)+
    theme_bw()+
  ggtitle("Breslow Depth > 0.5 mm & \u2264 1 mm") +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd2.clusts
```

```{r}

bd2.tsne$Risk <- factor(bd2.tsne$Risk, levels = c("High-Risk", "Low-Risk", "NS"))
bd2.risk <- ggplot(bd2.tsne, aes(x=tSNE1, y=tSNE2, color=Risk))+
  geom_point(size = 0.1)+
  ggtitle("Breslow Depth > 0.5 mm & \u2264 1 mm") +
    theme_bw()+
  scale_color_manual(values = c("#ef5350", "#26C6DA", "#B0BEC5")) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd2.risk
```

```{r}
logrank <- survdiff(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd2, Risk != "NS"))
logrank
```

```{r}
pval_label <- expression(paste(italic("p"), " < 0.0001"))
```


```{r}
fit.clust <- survfit(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd2, Risk != "NS"))

bd2_surv <- ggsurvplot(fit = fit.clust, data = subset(bd2, Risk != "NS"), censor.size = 1)$plot
bd2_surv <- bd2_surv +
  ggtitle("Breslow Depth > 0.5 mm & \u2264 1 mm") +
    theme_bw()+
    xlab("Months") +
    ylab("Survival Probability")+
  scale_color_manual(values = c("#ef5350", "#26C6DA")) +
  annotate(geom="text", x=25, y=0.2, label=pval_label,
              color="black", size = 8) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd2_surv
```

######### > 1mm and <= 4 mm

```{r}
set.seed(1)
all$Cluster <- as.factor(pr)
bd3 <- subset(all, Breslow_Depth > 1 & Breslow_Depth <= 4 & Stage == "Localized")

data2 <- data.table::rbindlist(lapply(1:25, calc_sig, bd3))
```

```{r}
bd3$Risk <- "NS"
bd3$Risk[bd3$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd3$Risk[bd3$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

bd3.tsne <- subset(tsne_vals,Breslow_Depth > 1 & Breslow_Depth <= 4)

bd3.tsne$Risk <- "NS"
bd3.tsne$Risk[bd3.tsne$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd3.tsne$Risk[bd3.tsne$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

```

```{r}
bd3.clusts <- ggplot(bd3.tsne, aes(x=tSNE1, y=tSNE2, color=Cluster))+
  geom_point(size = 0.1)+
    theme_bw()+
  ggtitle("Breslow Depth > 1 mm & \u2264 4 mm") +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd3.clusts
```

```{r}

bd3.tsne$Risk <- factor(bd3.tsne$Risk, levels = c("High-Risk", "Low-Risk", "NS"))
bd3.risk <- ggplot(bd3.tsne, aes(x=tSNE1, y=tSNE2, color=Risk))+
  geom_point(size = 0.1)+
  ggtitle("Breslow Depth > 1 mm & \u2264 4 mm") +
    theme_bw()+
  scale_color_manual(values = c("#ef5350", "#26C6DA", "#B0BEC5")) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd3.risk
```

```{r}
logrank <- survdiff(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd3, Risk != "NS"))
logrank
```

```{r}
pval_label <- expression(paste(italic("p"), " < 0.0001"))
```


```{r}
fit.clust <- survfit(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd3, Risk != "NS"))

bd3_surv <- ggsurvplot(fit = fit.clust, data = subset(bd3, Risk != "NS"), censor.size = 1)$plot
bd3_surv <- bd3_surv +
  ggtitle("Breslow Depth > 1 mm & \u2264 4 mm") +
    theme_bw()+
    xlab("Months") +
    ylab("Survival Probability")+
  scale_color_manual(values = c("#ef5350", "#26C6DA")) +
  annotate(geom="text", x=25, y=0.2, label=pval_label,
              color="black", size = 8) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd3_surv
```

######### > 4 mm

```{r}
set.seed(1)
all$Cluster <- as.factor(pr)
bd4 <- subset(all, Breslow_Depth > 4 & Stage == "Localized")

data2 <- data.table::rbindlist(lapply(1:25, calc_sig, bd4))
```

```{r}
bd4$Risk <- "NS"
bd4$Risk[bd4$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd4$Risk[bd4$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

bd4.tsne <- subset(tsne_vals,Breslow_Depth > 4)

bd4.tsne$Risk <- "NS"
bd4.tsne$Risk[bd4.tsne$Cluster %in% subset(data2, Risk == "HR" & Significance == "S")$Cluster] <- "High-Risk"
bd4.tsne$Risk[bd4.tsne$Cluster %in% subset(data2, Risk == "LR" & Significance == "S")$Cluster] <- "Low-Risk"

```

```{r}
bd4.clusts <- ggplot(bd4.tsne, aes(x=tSNE1, y=tSNE2, color=Cluster))+
  geom_point(size = 0.1)+
    theme_bw()+
  ggtitle("Breslow Depth > 4 mm") +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd4.clusts
```

```{r}

bd4.tsne$Risk <- factor(bd4.tsne$Risk, levels = c("High-Risk", "Low-Risk", "NS"))
bd4.risk <- ggplot(bd4.tsne, aes(x=tSNE1, y=tSNE2, color=Risk))+
  geom_point(size = 0.1)+
  ggtitle("Breslow Depth > 4 mm") +
    theme_bw()+
  scale_color_manual(values = c("#ef5350", "#26C6DA", "#B0BEC5")) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd4.risk
```

```{r}
logrank <- survdiff(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd4, Risk != "NS"))
logrank
```

```{r}
pval_label <- expression(paste(italic("p"), " < 0.0001"))
```


```{r}
fit.clust <- survfit(Surv(MSS_Time, MSS) ~ Risk, data = subset(bd4, Risk != "NS"))

bd4_surv <- ggsurvplot(fit = fit.clust, data = subset(bd4, Risk != "NS"), censor.size = 1)$plot
bd4_surv <- bd4_surv +
  ggtitle("Breslow Depth > 4 mm") +
    theme_bw()+
    xlab("Months") +
    ylab("Survival Probability")+
  scale_color_manual(values = c("#ef5350", "#26C6DA")) +
  annotate(geom="text", x=25, y=0.2, label=pval_label,
              color="black", size = 8) +
  theme(legend.position = "none", plot.title=element_text(hjust=0.5, family = "serif", size = 24), axis.title = element_text(family = "serif", size = 24), axis.text = element_text(family = "serif", size = 22))
bd4_surv
```


```{r}
final_plot <- ggpubr::ggarrange(bd1_surv, bd1.risk,
                                bd2_surv, bd2.risk,
                                bd3_surv, bd3.risk,
                                bd4_surv, bd4.risk, nrow = 4, ncol = 2, common.legend = TRUE, legend = "bottom")
### 1600 x 2000 saved jpeg
```
